Bringing the Power of eBPF to Open vSwitch

### 2018-10-25 paper.tex
Edit 4.4.2 - 4.4.4

\subsubsection{Memory Management and Locks}\label{memorym}
For every stage of optimizations, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the umempool APIs we introduce, as well as at the
packet metadata allocation, struct dp\_packet, for every received packet.
We introduce two major APIs, umem\_elem\_get(), which gets N free element from the
umempool, and umem\_elem\_put(), which places back the free umem buffer to
the umempool, similar to the concept of producer (put) and consumer (get).
We implemented three different data structures of umempool as below
to compare their performance.
\begin{itemize}
\item FILO-list\_head: We reversed first 8-byte of the umem buffer as a next pointer,
pointing to the next available element in umempool, and implement FIFO access get/set
APIs. The list\_head pointer always points to the first available umem element.
\item FIFO-ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region.
Consumer takes elements from the tail pointer and when placing back, producer, puts
them into where head pointer points to.
\item FILO-ptr\_array: Similar to Linux kernel's skb\_array~\cite{skbarray},
an array of elements are allocated
in continuous memory region, but both consumer and producer take elements
from beginning of the array.
\end{itemize}

Although we allocate one umem per netdev, there might be multiple queues per
netdev sharing the same umem. As a result, the above three data structures
require a mutex lock before accessing umem freelist. Linux perf reports
pthread mutex lock related APIs as ones of the top CPU utilization functions.
We change our design by 1) allocating per-queue umem region and 2)
allocating one PMD thread per queue. As a result, no lock is needed because
each queue has only one thread and its own set of umem elements.

\subsubsection{Metadata Allocation}
Moving forward, Linux perf shows that the packet metadata allocation
takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
So instead of allocating packet metadata at packet reception time, we pre-allocate
the packet metadata and implement two data structures to compare their performances:
\begin{itemize}
\item Embedding in packet buffer: Since we already allocate 2K chunk for each
umem packet buffer, we reserve the first 256-byte in each umem element as struct dp\_packet
and initialize the dp\_packet as much as we can at allocation time.
This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
is used for both packet data and metadata.
\item Separate from packet buffer: This design simply allocates a continuous memory region
storing an array of packet metadata, and initialize them as much as we can.
\end{itemize}

@@

With the above design change, we found that using the FILO-ptr\_array in section~\ref{memorym} with
separate from packet buffer design yields the best performance. Largely due to both data structures
have better spatial locality and more batching friendly, when all data are allocated together.
For example, setting up 32 packet metadatas in an array definitely incurs
less cache misses than setting up in 32 elements in umem.
And getting 32 free umem elements from ptr\_array has similar benefits.
With the above decision, the OVS rxdrop shows around $19Mpps$, closed
to the baseline xdpsock rxdrop performance.

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
We continued measuring the performance of OVS-AFXDP l2fwd and
observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock-l2fwd.
We found that the OVS PMD thread under rxdrop
has much fewer context switches compared to the l2fwd, indicating that
the PMD process spends much more time in kernel space than in userspace.
By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
the success of send (step 4) immediately after issusing send (step 3).
We change the design by calling sendmsg syscall (step 3) only when TX ring is
closed to full, e.g., when 3/4 ring elements have been used.
In specific case, instead of issuing sendmsg syscall for a batch of 32 packets
and making sure they are done, we only issue send when there are 512 outstanding
packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.

>>>>

\subsubsection{Memory Management and Locks}\label{memorym}
For every stage of optimizations, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above PMD netdev optimization, we observe the
new bottleneck is the umempool APIs that we introduced.
We introduce two major umempool APIs, umem\_elem\_get(), which gets N free
element from the umempool, and umem\_elem\_put(), which places back the free
umem buffer to the umempool. We implement three different data structures for
umempool management as below.
\begin{itemize}
\item LIFO-list\_head: We reverse the first 8-byte of a umem chunk as a next
pointer that pointing to the next available element in umempool, and implement
LIFO (last in, first out) access get/put APIs. In this design, we maintain the
list\_head pointer that always points to the first available umem element,
and we get and put elements from list\_head.
\item FIFO-ptr\_ring: This design maintains an extra FIFO (first in, first out)
pointer ring that is similar to Linux kernel's ptr\_ring. The ring is an array
of elements that allocated in a continuous memory region, where each element
points to an available umem chunk. We keep track of head and tail pointer for
the FIFO-ptr\_ring. The consumer gets elements from the tail pointer,
and then the producer puts elements back to where head pointer points to.
\item LIFO-ptr\_array: This design is similar to Linux kernel's
skb\_array~\cite{skbarray} where an array of pointers are allocated
in contiguous memory region. Each element in the array points to an available
memory chunk in umem.  We keep track of a head pointer to the array,
and both consumer and producer get/put elements from the head pointer.
\end{itemize}

Initially, we allocate one umem per netdev. Since there might be multiple
queues per netdev sharing the same umem, the above three data structures
require a mutex lock to protect umempool accessing. In this design, Linux
perf reports pthread mutex lock related APIs as one of the top CPU
utilization function.
We then change our design by 1) allocating per-queue umem region and 2)
allocating one PMD thread per queue. As a result, no lock is needed because
each queue has only one thread and its own set of umem elements.

\subsubsection{Metadata Allocation}
Moving forward, Linux perf shows that the packet metadata allocation
takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
So instead of allocating packet metadata at packet reception time, we pre-allocate
the packet metadata and implement two data structures to compare their performance:
\begin{itemize}
\item Embedding in packet buffer: As we already allocate 2KB chunk for each
umem packet buffer, we reserve the first 256-byte in each umem chunk as struct dp\_packet
and initialize the dp\_packet's packet independent fields  at allocation time.
This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
is used for both packet data and metadata.
\item Separate from packet buffer: This design allocates a contiguous memory
region storing an array of packet metadata, and initialize their packet
independent fields.
\end{itemize}

@@

With the above design change, we find that using the LIFO-ptr\_array in
Section~\ref{memorym} with separated packet metadata allocation yields the
best performance. It is because both data structures have better spatial
locality and there are more batching friendly.
For example, accessing 32 packet metadata in an array incurs less cache
misses than accessing the packet metadata 32 umem chunks.
The ptr\_array for umempool management outperforms the other two designs
for the similar reason. With the above design decisions, the OVS rxdrop
can achieve similar performance as the baseline xdpsock rxdrop at around
$19Mpps$.

\subsubsection{Batching Send Syscall}
%574.790 M/sec  
With all the aforementioned optimizations, We continue measuring the
performance of OVS-AFXDP l2fwd and observe only $4Mpps$, compared to
$17Mpps$ baseline xdpsock-l2fwd.
We find that the OVS PMD thread under rxdrop has much fewer context
switches compared to the l2fwd, indicating that
the PMD process spends much more time in kernel space than in userspace.
By using strace, we find that the OVS-AFXDP l2fwd experiment calls sendto
system call at very high rate. It is because we design to check the
completion of send immediately after issusing send as shown in step 3 and
4 in Figure~\ref{afxdptx}.
We modify the design by calling sendmsg syscall (step 3) only when TX ring is
close to full, e.g., when 3/4 ring elements have been used.
For example, instead of issuing sendmsg syscall for a batch of 32 packets
and making sure they are finished, we only issue send when there are 512
outstanding packets.  With this change, the OVS-AFXDP l2fwd experiment
can achieve around $14Mpps$.
